Examining Whether Approximate Visualizations Impact Perception of Corrected Visualizations

### 2016-12-08 paper/paper.tex
More details on the study and new figures

% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps (small caps for compsoc).
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by the IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{A}{s} data science grows in interest and importance, data analysts want to derive insights from increasingly large datasets.
Depending on the dataset, it could take a system seconds or minutes to construct accurate summary visualizations of the data.
The time to generate these visualizations drastically impedes the process of exploratory data visualization, where analysts often consider many visualizations in sequence.
A common technique to enable exploratory analysis of large datasets is to sample the data for immediate consumption \cite{moritz2017pangloss}.
Precise results considering the entire dataset are either updated progressively \cite{fisher2012trust} or asynchronously \cite{moritz2017pangloss, kay2016ish}.

@@

Through an X-person experiment on Amazon Mechanical Turk, we found \todo{write when we have data}.

@@

\begin{figure*}[!ht]
\centering
\subfloat[Instructions for the study in general.]{\includegraphics[width=0.18\textwidth]{1.png}%
\label{fig:one}}
\hfil
\subfloat[First approximate visualization (here airlines).]{\includegraphics[width=0.18\textwidth]{2.png}%
\label{fig:two}}
\hfil
\subfloat[Questions about first approximate visualization.]{\includegraphics[width=0.18\textwidth]{3.png}%
\label{fig:three}}
\hfil
\subfloat[Second approximate visualization (here states).]{\includegraphics[width=0.18\textwidth]{4.png}%
\label{fig:four}}
\hfil
\subfloat[Questions about second approximate visualization.]{\includegraphics[width=0.18\textwidth]{5.png}%
\label{fig:five}}
\hfil
\subfloat[Instructions for the precise visualizations.]{\includegraphics[width=0.18\textwidth]{6.png}%
\label{fig:six}}
\hfil
\subfloat[First precise visualization (here airlines and difference visualization).]{\includegraphics[width=0.18\textwidth]{7.png}%
\label{fig:seven}}
\hfil
\subfloat[Second precise visualization (here states and difference visualization).]{\includegraphics[width=0.18\textwidth]{8.png}%
\label{fig:eight}}
\hfil
\subfloat[Questions about both precise visualizations.]{\includegraphics[width=0.18\textwidth]{9.png}%
\label{fig:nine}}
\hfil

@@

We sought to create an experiment which mimicked the experience of Exploratory Data Analysis with an approximate query system while still maintaining control over participant experiences.
In approximate query systems, users often internalize approximate visualizations, using the gained knowledge to inform the next visualization to create.
Systems, such as Pangloss \cite{moritz2017pangloss}, can later generate precise visualizations based on the data.
However, the user may have moved on to investigate a new visualization.

Our study flow (\figref{fig:example}) aimed to mimic this experience. %TODO: add; probably include a few screenshots about what participants actually saw.
Participants viewed two approximate visualizations followed by two precise visualizations.
The study enforced participants view the visualizations for at least 10 seconds before advancing. %TODO: why do we pick 10 seconds?
To ensure that participants had considered the approximate data, we asked a few questions after viewing each approximate visualization informed by Brehmer and Munzner's work \cite{brehmer2013multi}.
Answering these questions simulated how a data analyst would consider and reflect on their data.
The questions also provided us a baseline level of comprehension of visualizations of our participants.
After viewing both precise visualizations, participants answered a larger set of questions based on Brehmer and Munzner's task taxonomy \cite{brehmer2013multi}.
Participants were asked to rate their confidence in their answer to each question in both sets on a sliding scale from ``I don't remember'' to ``I'm certain''.
Participants were informed that the first two visualizations showed approximations of their data, allowing us to mimic the experience of looking at an approximate query result.
%Do we want to go into more detail on the questions? In a 'dependent variables' section?

@@

The two visualizations participants saw were based on the same dataset about flights in the United States \footnote{\url{http://datahub.io/dataset/open-flights}}.
\figref{fig:types} shows visualizations of approximate and precise data.
Both visualizations were bar graphs, which are common in approximate query systems \cite{moritz2017pangloss}.
The y-axis varied between the two visualizations, showing either the state the flight left from or the airline the flight was operated by.
The bars were sorted in descending order so that the state or airline with the most flights is at the top.
The x-axis in both visualization was a count value.
We chose horizontal bar charts so that long labels are easier to read.
To visually separate visualizations of approximate and precise data, we colored the bars approximate visualizations orange (\figref{fig:approx}) and precise visualizations blue (\figref{fig:precise}).

We used one month of flight data to generate the visualizations.
The precise visualization of number of flights per state uses data where the airline was ``JetBlue'' and the visualization of the flight per airline uses data from ``New York State''.
In the study, we randomized the airline and state to focus on.
The approximate visualizations use a random subset of the data for the precise visualizations with replacement.
We then computed the approximate values by dividing the count values by the fraction of rows used in the sample.

Participants were recruited from Amazon Mechanical Turk (AMT), a common tool to run perception studies for visualizations~\cite{heer2010crowdsourcing}.
Similar to other studies of human perception of visualizations, we restricted participation to people on AMT with at least 1,000 tasks completed and an approval rate above 95\%.
Participation was restricted to the United States to ensure familiarity with the airlines and states in the dataset.
Participants were compensated \$2.00 for completing the study.

@@

%Probably make a figure for this too.
We evaluated three techniques for presenting the precise visualization.
Our control condition presented the precise data as both the approximate and precise data.
We further tested presenting the approximate and precise visualizations side-by-side, and a visualization highlighting the difference to approximate data on the precise visualization (\figref{fig:diff}).
The difference visualization aims to highlight the errors of the approximation as done in Pangloss \cite{moritz2017pangloss}.

@@

We randomized visualization order, approximate sample, focus state, focus airline, and study condition for each participant.
As previously stated, the variation level of the approximate visualizations varied by visualization.

>>>>

% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps (small caps for compsoc).
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by the IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{A}{s} data science grows in interest and importance, data analysts want to derive insights from increasingly large datasets.
Depending on the dataset, it could take a system seconds or minutes to construct accurate summary visualizations of the data.
The time to generate these visualizations drastically impedes the process of exploratory data visualization \cite{liu2014effects}, where analysts often consider many visualizations in sequence.
A common technique to enable exploratory analysis of large datasets is to sample the data for immediate consumption \cite{moritz2017pangloss}.
Precise results considering the entire dataset are either updated progressively \cite{fisher2012trust} or asynchronously \cite{moritz2017pangloss, kay2016ish}.

@@

Through an 38-person experiment on Amazon Mechanical Turk, we found that exposure to an incorrect visualization biases the recollection of data from the precise visualization.
However, we did not see enough evidence to support the hypothesis that highlighting the differences reduces this bias.

@@

\begin{figure*}[!ht]
\centering
\subfloat[Instructions for the study in general.]{\includegraphics[width=0.18\textwidth]{1.png}%
\label{fig:one}}
\hfil
\subfloat[First approximate visualization (here airlines).]{\includegraphics[width=0.18\textwidth]{2.png}%
\label{fig:two}}
\hfil
\subfloat[Questions about first approximate visualization.]{\includegraphics[width=0.18\textwidth]{3.png}%
\label{fig:three}}
\hfil
\subfloat[Second approximate visualization (here states).]{\includegraphics[width=0.18\textwidth]{4.png}%
\label{fig:four}}
\hfil
\subfloat[Questions about second approximate visualization.]{\includegraphics[width=0.18\textwidth]{5.png}%
\label{fig:five}}
\hfil
\subfloat[Instructions for the precise visualizations.]{\includegraphics[width=0.18\textwidth]{6.png}%
\label{fig:six}}
\hfil
\subfloat[First precise visualization (here airlines and difference visualization).]{\includegraphics[width=0.18\textwidth]{7.png}%
\label{fig:seven}}
\hfil
\subfloat[Second precise visualization (here states and difference visualization).]{\includegraphics[width=0.18\textwidth]{8.png}%
\label{fig:eight}}
\hfil
\subfloat[Questions about both precise visualizations.]{\includegraphics[width=0.18\textwidth]{9.png}%
\label{fig:nine}}
\hfil
\subfloat[Questions about the demographics.]{\includegraphics[width=0.18\textwidth]{10.png}%
\label{fig:ten}}

@@

We sought to create an experiment which mimicked the experience of Exploratory Data Analysis with an approximate query system while still maintaining control over participant experiences.
In approximate query systems, users often internalize approximate visualizations, using the gained knowledge to inform the next visualization to create.
Systems, such as Pangloss \cite{moritz2017pangloss}, can later generate visualizations of the precise results.
However, the user may have moved on to investigate a new visualization before the precise result is computed.

\subsubsection{Study Flow}

Our study flow (\figref{fig:example}) aimed to mimic the experience of a data scientist analyzing a large dataset about flights in an approximate query processing system.
Participants viewed two approximate visualizations followed by two precise visualizations and then asked questions about the data in them.
Before we showed the first visualization, we instructed each participant to focus on a particular airline and state.
This mimics a data analysis session where analysts form hypotheses that affect what they focus on rather than trying to look at every possible aspect (e.g. airlines, states) with equal attention.
The study enforced participants view the visualizations for at least 10 seconds before advancing. %TODO: why do we pick 10 seconds?
To ensure that participants had considered the approximate data, we asked a few questions after viewing each approximate visualization informed by Brehmer and Munzner's work \cite{brehmer2013multi}.
Answering these questions simulated how a data analyst would consider and reflect on their data.
The questions also provided us a baseline level of comprehension of visualizations of our participants.
After viewing both precise visualizations, participants answered a larger set of questions similar to the ones we asked about the visualization of the approximate data.
Participants were asked to rate their confidence in their answer to each question in both sets on a sliding scale from ``I don't remember'' to ``I'm certain''.
Participants were informed that the first two visualizations showed approximations of their data, allowing us to mimic the experience of looking at an approximate query result.

\subsubsection{Visualizations in the Study}\label{sec:vizes}

The visualizations participants saw were based on the same dataset about flights in the United States \footnote{\url{http://datahub.io/dataset/open-flights}}.
All visualizations were bar graphs, which are common in approximate query systems \cite{moritz2017pangloss}.
The y-axis varied between the two visualizations, showing either the state the flight left from or the airline the flight was operated by.
The length of the bar encodes the number of flights operated by a particular airline or the number of flights out of different states.
The bars were sorted in descending order so that the state or airline with the most flights is at the top.
We chose horizontal bar charts so that long labels are easier to read.
To visually separate visualizations of approximate and precise data, we colored the bars approximate visualizations orange (\figref{fig:approx}) and precise visualizations blue (\figref{fig:precise}).

We used one month of flight data to generate the visualizations.
The precise visualization of number of flights per state uses data where the airline was ``JetBlue'' and the visualization of the flight per airline uses data from ``New York State''.
The approximate visualizations use a random subset of the data for the precise visualizations with replacement.
We then computed the approximate values by dividing the count values by the fraction of rows used in the sample.

@@

We designed three different visualizations of the precise data to understand whether they affect the bias (RQ2).
The baseline condition was to show only the precise data (\figref{fig:precise}).
We could also show participants both the approximate (\figref{fig:approx}) and precise (\figref{fig:precise}) visualization side by side.
This method reminds participants of what they have previously seen so that they can clear their memory from approximate results.
Lastly, we could overlay the approximate data as small ticks on the bar chart that shows the precise data (\figref{fig:diff}).

\subsubsection{Participants}

We recruited 41 participants from Amazon Mechanical Turk (AMT), a common tool to run perception studies for visualizations~\cite{heer2010crowdsourcing}.
Similar to other studies of human perception of visualizations, we restricted participation to people on AMT with at least 1,000 tasks completed and an approval rate above 95\%.
Participation was restricted to the United States to ensure familiarity with the airlines and states in the dataset.
Participants were compensated \$2.00 for completing the study.

@@

In our study, each participant was asked to focus on a particular airline and state.
All questions were about the one airline and state directly or comparisons with the selected airline or state.
The state and airline to focus on were picked at random from the top three airlines and states and remained the same throughout the study.

%Probably make a figure for this too.
We evaluated three techniques for presenting the precise visualization (\secref{sec:vizes}).
Our control condition presented the precise data as both the approximate and precise data.
We further tested presenting the approximate and precise visualizations side-by-side, and a visualization highlighting the difference to approximate data on the precise visualization (\figref{fig:diff}).
The difference visualization aims to highlight the errors of the approximation as done in Pangloss \cite{moritz2017pangloss}.

@@

In this study, we randomized the order that participants saw the two visualizations in; some saw the visualization for airlines first, others saw the visualization of states first.
We used the same order for the approximate and precise visualizations.
We also randomized approximate sample, focus state, focus airline, and study condition for each participant.
As previously stated, the sample level of the approximate visualizations varied by visualization.
